Efficient static analysis of Marlowe contracts

### 2020-07-06 papers/static_analysis2020/paper.tex
SCP-843 - Address reviews from ISOLA 2020

The techniques described helped us reduce the analysis time of an implementation
that followed the semantics closely and took a couple of minutes to analyse contracts
of a few kilobytes, to one where the same contract takes less
than a second to analyse, and where a four-person crowdfunding contract that
fully expanded occupies about 19 megabytes, can be analysed in
around 10 minutes. 
We have classified optimisation techniques as lightweight and heavyweight.

@@

\subsection{Overview}

@@

High level libraries like SBV, and even standards like SMTLib, support
the use and construction of high level abstractions,
e.g: custom data-types, list, sets\dots{} Unfortunately, even though
high level abstractions aid reasoning about code, they often
prevent optimisations, since they abstract out aspects that
in our particular case may be concrete.

@@

In our experiments, the heavyweight optimisations considerably reduced
the requirements of both processing time and memory for Marlowe's static
analysis. %In Tables~\ref{table1},~\ref{table2},~\ref{table3},~and~\ref{table4},
We present below the results of measuring the performance of static analysis on
four example  contracts,
before\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemantics.hs} [last visited 19-05-2020]}
and after\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemanticsFastVerbose.hs} [last visited 19-05-2020]}
the heavyweight optimisations.

>>>>

The techniques described helped us reduce the analysis time of an implementation
that followed the semantics closely and took a couple of minutes to analyse contracts
with a source code of a few kilobytes, to one where the same contract takes less
than a second to analyse, and where a four-person crowdfunding contract that
fully expanded occupies about 19 megabytes, can be analysed in
around 10 minutes. 
We have classified optimisation techniques as lightweight and heavyweight.

@@

@@

High level libraries like SBV, and even standards like SMTLib, support
the use and construction of high level abstractions,
e.g: custom data-types, list, sets\ldots{} Unfortunately, even though
high level abstractions aid reasoning about code, they often
prevent optimisations, since they abstract out aspects that
in our particular case may be concrete.

@@

In our experiments, the heavyweight optimisations considerably reduced
the requirements of both processing time and memory for Marlowe's static
analysis. %In Tables~\ref{table1},~\ref{table2},~\ref{table3},~and~\ref{table4},
We present below the results of measuring the performance of static analysis on
four example contracts,
before\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemantics.hs} [last visited 19-05-2020]}
and after\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemanticsFastVerbose.hs} [last visited 19-05-2020]}
the heavyweight optimisations. We could not obtain data for some of the experiments because they
required more RAM than there was available or too much processing time.
